---
title: "Chapter 2"
tutorial:
  id: "code.r-journalism/chapter-2"
  version: .9
output:
  learnr::tutorial:
    progressive: true
    theme: lumen
    highlight: espresso
    include:
      before_body: _navbar.html
runtime: shiny_prerendered
---


```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
knitr::opts_chunk$set(echo = FALSE)
burgers <- data.frame(id=c(60006,60007,60008,60009, 60010), name=c("Bob", "Linda", "Louise", "Tina", "Gene"), age=c(45, 44, 12, 13, 11), grade=c(NA, NA, "seventh", "eighth", "sixth"))

burgers$grade <- factor(burgers$grade)

burgers$id <- factor(burgers$id)
burgers$name <- as.character(burgers$name)
burgers <- rename(burgers, `First name`=name)

burgers2 <- data.frame(id=c(60006,60007,60008,60009, 60010), name=c("Bob", "Linda", "Louise", "Tina", "Gene"), age=c(45, 44, 12, 13, 11), grade=c(NA, NA, "seventh", "eighth", "sixth"))
burgers2$grade <- factor(burgers2$grade)

burgers2$id <- factor(burgers2$id)
burgers2$name <- as.character(burgers2$name)
burgers2 <- rename(burgers2, employee_id=id)


nospace <- function(a) {
  a <- gsub(" ","",a) 
  gsub("\\n", "", a)
  gsub("\\`", "", a)
}


is_bad_code <- function(user_code, check_code, envir_result) {

  is_null <- is.null(eval(parse(text = user_code)))
  e_rows <- nrow(eval(parse(text = user_code)))
  r_rows <- nrow(eval(parse(text = check_code), envir = envir_result))
  e_cols <- ncol(eval(parse(text = user_code)))
  r_cols <- ncol(eval(parse(text = check_code), envir = envir_result))
  e_type <- typeof(eval(parse(text = user_code)))
  r_type <- typeof(eval(parse(text = check_code), envir=envir_result))
  e_len <- length(eval(parse(text = user_code)))
  r_len <- length(eval(parse(text = check_code)))
  if (is_null!=T){
    if (e_len!=0 & r_len!=0) {
      if (e_type==r_type) {
        if (e_type!="character" & e_type!="double" & e_type!="integer") {
          if (e_rows==r_rows && e_cols==r_cols) {
            eval(parse(text = user_code)) != eval(parse(text = check_code), envir = envir_result)
          } else {
            TRUE
          }
        } else { eval(parse(text = user_code)) != eval(parse(text = check_code), envir = envir_result) }
      } else {
        TRUE
      }
    } else { TRUE }
  } else { nospace(user_code) != nospace(check_code)}
}

checker <- function(label, user_code, check_code, envir_result, evaluate_result, ...) {
  if (is_bad_code(user_code, check_code, envir_result)) {
    return(list(message = "Code wasn't right!", correct = FALSE, location = "append"))
  } else {
    return(list(message = "Great Job!", correct = TRUE, location = "append"))
  }
}  

tutorial_options(exercise.timelimit = 30, exercise.checker = checker)

```


## CSVs

```{r quiz1, echo=FALSE}
quiz(caption = "Quiz 1",
  question("If you import a CSV with read.csv(), what argument do you pass to make sure strings are not interpreted as factors?",
    answer('read.csv("data.csv", stringsAsFactors=F)', correct=T),
    answer('read.csv("data.csv", interpretStrings=T)'),
    answer('read.csv("data.csv", factors=F)'),
    answer('read.csv("data.csv", strings=T)'),
    random_answer_order= TRUE
  ),
  question("When exporting a csv, how do you get rid of NAs with the readr package?",
    answer('write_csv(df_csv, "data/transformed_data.csv", na="")', correct = TRUE),
    answer('write.csv(df_csv, "data/transformed_data.csv", na="")'),
    answer('write_csv(df_csv, "data/transformed_data.csv", NA="")'),
    answer('write.csv(df_csv, "data/transformed_data.csv", is.na=F)'),
    random_answer_order= TRUE
  )
)

```

## Excel

```{r quiz2, echo=FALSE}
quiz(caption = "Quiz 2",
  question("If you import an Excel file and it has 5 buffer rows at the top before the real data starts, how do you deal with that using the **readxl** package?",
    answer('read_excel("stats.xls", sheet=1, skip=5)', correct=T),
    answer('read_excel("stats.xls", sheet=1, row=5)'),
    answer('read_excel("stats.xls", sheet=1, row=6)'),
    answer('read_excel("stats.xls", sheet=1, row=6)'),
    random_answer_order= TRUE
  ),
  question("When exporting a csv, how do you get rid of NAs with the readr package?",
    answer('write_csv(df_csv, "data/transformed_data.csv", na="")', correct = TRUE),
    answer('write.csv(df_csv, "data/transformed_data.csv", na="")'),
    answer('write_csv(df_csv, "data/transformed_data.csv", NA="")'),
    answer('write.csv(df_csv, "data/transformed_data.csv", is.na=F)'),
    random_answer_order= TRUE
  )
)
```


### Column names

What if you import in a spreadsheet with a space in the column name like below?

```{r show_burgers}
burgers
```

How do you refer to the the *First name* variable/column with **$** and not the **[]**?

```{r v2, exercise=TRUE}
# Add to the line below 

burgers$
```


```{r v2-check}
burgers$`First name`
``` 



### Clean up column name

Let's create a new data frame **burgers2** from **burgers** because we're going to change it up.

Run the command on the *First name* variable/column to strip out the spaces and characters and replace them with periods.


```{r v3, exercise=TRUE}
# Modify the line of code below

burgers2 <- burgers

colnames(burgers2) <- _______(colnames(burgers2))

burgers2
```


```{r v3-check}
burgers2 <- burgers

colnames(burgers2) <- make.names(colnames(burgers2))

burgers2
``` 


### Rename a column

Use the special **dplyr** function to rename the *id* variable in **burgers** to *employee_id*



```{r rn, exercise=TRUE}
# Modify the line of code below

burgers2 <- 
  
burgers2
```


```{r rn-check}
burgers2 <- rename(burgers, employee_id=id)

burgers2
``` 



### Filter out NAs

```{r show_burgers2}
burgers2
```

Create a new data frame called **kids** by filtering out the NAs from *grade* column in **burgers**.


```{r na, exercise=TRUE}
# Modify the line of code below 

kids <- _______(burgers2, !____(______))

kids
```


```{r na-check}
kids <- filter(burgers2, !is.na(grade))

kids
``` 

