---
title: "Chapter 3"
tutorial:
  id: "code.r-journalism/chapter-3"
  version: .85
output:
  learnr::tutorial:
    progressive: true
    theme: lumen
    highlight: espresso
    include:
      before_body: _navbar.html
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(lubridate)
library(stringr)
knitr::opts_chunk$set(echo = FALSE)
bo <- readRDS("bw.rds")
colnames(bo) <- make.names(colnames(bo))
orcas <- bo %>% filter(Species=="Killer Whale; Orca")
orcas2 <- select(orcas, -Species)
orcas3 <- orcas2 %>% arrange(Facility)
orcas4 <- mutate(orcas3, status_num=case_when(Status=="DEAD" ~ 1, TRUE ~ 0))
alive_dead <- orcas4 %>%
  summarize(total=n_distinct(NOAA.ID), died=sum(status_num))

perc_dead <- alive_dead %>%
  mutate(percent_dead=died/total*100)

owner_percent <- bo %>%
   group_by(Owner, Species) %>%
   summarize(total=n()) %>%
   mutate(percent=round(total/sum(total, na.rm=T)*100,2)) %>%
   select(-total) %>%
   spread(Species, percent)

owner_total <- bo %>%
   group_by(Owner, Species) %>%
   summarize(total=n())

bo2 <- mutate(bo, Name=str_to_lower(Name),
              Sex=str_to_title(Sex))

bo3 <- mutate(bo2,
             sex_species=str_c(Sex, " ", Species))

bo4 <- mutate(bo3,
             sex_species=str_replace(sex_species, " Killer Whale;", "")) %>%
      mutate(sex_species=str_replace(sex_species, " Whale", ""))

bo_dates <- bo %>%
  mutate(Captivity.Date=dmy(Captivity.Date))

bo_dates2 <- bo_dates %>%
  mutate(month_number=month(Captivity.Date))
bo_dates3 <- bo_dates2  %>%
  mutate(month_label=month(Captivity.Date, label=T))


nospace <- function(a) {
  b <- gsub(" ","",a) 
  gsub("\\n", "", b)
}



is_bad_code <- function(user_code, check_code, envir_result) {

  is_null <- is.null(eval(parse(text = user_code)))
  e_rows <- nrow(eval(parse(text = user_code)))
  r_rows <- nrow(eval(parse(text = check_code), envir = envir_result))
  e_cols <- ncol(eval(parse(text = user_code)))
  r_cols <- ncol(eval(parse(text = check_code), envir = envir_result))
  e_type <- typeof(eval(parse(text = user_code)))
  r_type <- typeof(eval(parse(text = check_code), envir=envir_result))
  e_len <- length(eval(parse(text = user_code)))
  r_len <- length(eval(parse(text = check_code)))
  if (is_null!=T){
    if (e_len!=0 & r_len!=0) {
      if (e_type==r_type) {
        if (e_type!="character" & e_type!="double" & e_type!="integer" & e_type!="logical") {
          if (e_rows==r_rows && e_cols==r_cols) {
            eval(parse(text = user_code)) != eval(parse(text = check_code), envir = envir_result)
          } else {
            TRUE
          }
        } else { eval(parse(text = user_code)) != eval(parse(text = check_code), envir = envir_result) }
      } else {
        TRUE
      }
    } else { TRUE }
  } else { nospace(user_code) != nospace(check_code)}
}

checker <- function(label, user_code, check_code, envir_result, evaluate_result, ...) {
  if (is_bad_code(user_code, check_code, envir_result)) {
    return(list(message = "Code wasn't right!", correct = FALSE, location = "append"))
  } else {
    return(list(message = "Great Job!", correct = TRUE, location = "append"))
  }
}  

tutorial_options(exercise.timelimit = 30, exercise.checker = checker)
#tutorial_options(exercise.checker = checkthat::check_exercise)


```

## Transforming and analyzing data

Consider this Marine Mammal Inventory Report data set from the National Oceanic and Atmospheric Administration.

It has the location of every Beluga and Orca whale in the United States since the '70s.

Currently, the data frame is in the environment and has 494 rows.

This view shows the first six rows.

```{r data_view}
head(bo)
```

Each animal is represented by a *NOAA.ID* (because some do not have names) and might show up multiple times in the data frame.

For example, the Beluga Whale Blanchon is NOA0000719 and has been at three facilities but his latest *Status* shows him as "DEAD".

### Filter

Please create a data frame of Orca whales only from the **bo** data frame.

We'll name the new data frame **orcas**.


```{r f, exercise=TRUE}
# Add to the line below 

orcas <- bo %>%
  
unique(orcas$Species)
```


```{r f-check}
orcas <- bo %>% filter(Species=="Killer Whale; Orca")
  
unique(orcas$Species)
``` 


<div id="f-hint">
**Hint:** Delete the commented line out before submitting.
</div>

### Select

Drop the column/variable *Species* since there's only one variable in it now.

```{r s, exercise=TRUE}
# Add to the line below

orcas2 <- orcas
  
colnames(orcas2)
```


```{r s-check}
orcas2 <- select(orcas, -Species)

colnames(orcas2)
``` 


### Arrange

Sort the **orcas** data frame based on the *Facility* column in alphabetical order.


```{r a, exercise=TRUE}
# Add to the line below

orcas3 <- orcas2
  

head(orcas3)
```


```{r a-check}
orcas3 <- orcas2 %>% arrange(Facility)
  

head(orcas3)
``` 


### Mutate

Create a new column called *status_num* to fill with a 1 if *Status* is "DEAD" or 0 if alive. Save the transformed data frame back into **orcas**


```{r m, exercise=TRUE}
# Transform the code below

orcas4 <- orcas3
  

head(orcas4)
```


```{r m-check}
orcas4 <- mutate(orcas3, status_num=case_when(Status=="DEAD" ~ 1, TRUE ~ 0))
  

head(orcas4)
``` 

<div id="m-hint">
**Hint:** You may want to use the dplyr `case_when()` function nested in another function.
</div>

### Summarize


Count up the number of total Orcas by counting the unique *NOAA.ID*s and how many have died by adding up the *status_num*. You want to summarize with two columns: *total* (by counting distinct NOAA.IDS and adding up number of those that have *died*)



```{r sum, exercise=TRUE}
# Add to the line below

alive_dead <- orcas4
  
alive_dead
```


```{r sum-check}
alive_dead <- orcas4 %>%
  summarize(total=n_distinct(NOAA.ID), died=sum(status_num))
``` 

<div id="sum-hint">
**Hint:** You may want to use the dplyr `n_distinct()` function nested in another function.
</div>

### Percent


What's the percent of dead Orcas?

Create a new column/variable called *percent_dead*.

You do **not** need to round the results.


```{r m2, exercise=TRUE}
# Add to the line below

perc_dead <- alive_dead

perc_dead
```


```{r m2-check}
perc_dead <- alive_dead %>%
  mutate(percent_dead=died/total*100)

perc_dead
``` 


<div id="m2-hint">
**Hint:** But don't forget to multiply the results by 100 to get the percent.
</div>

Now figure how the percent of Orcas still alive.

Create a new column/variable called *percent_alive*.

You do **not** need to round the results.

```{r m3, exercise=TRUE}
# Add to the line below

percent_dead_alive <- perc_dead

percent_dead_alive
```


```{r m3-check}
percent_dead_alive <- perc_dead %>%
  mutate(percent_alive=100-percent_dead)

percent_dead_alive
``` 


### Group By

Consider the **bo** data frame again.


```{r data_view2}
head(bo)
```

Which *Facility* had the most Orcas and Belugas combined? 

Go ahead and call the new column *mammal_count*.

List them from highest count to lowest.


```{r fc, exercise=TRUE}
# Add to the line below

facilities_count <- bo

facilities_count
```


```{r fc-check}
facilities_count <- bo %>%
  group_by(Facility) %>%
  summarize(mammal_count=n()) %>%
  arrange(-mammal_count)

facilities_count
``` 



<div id="fc-hint">
**Hint:** `n()` is used to count. But use it correctly.
</div>

## Tidying and joining data

Consider the **bo** data set again.

```{r data_view3}
head(bo)
```

### Turn a tall data frame wide

Summarize the data and count the total number of Orcas and Belugas (separately) per facility from the *Species* variable. Use the *mammal_count* as a column name again.

Then turn the tall data into a wide data frame that has these columns: Facility, Belugas, Orcas.

```{r w, exercise=TRUE}
# Add to the line below

facilities_count <- bo %>%
   group_by(Facility, ________) %>%
   summarize(_______=___) %>%
   ______(_______, _____)

facilities_count
```


```{r w-check}
facilities_count <- bo %>%
  group_by(Facility, Species) %>%
  summarize(mammal_count=n()) %>%
  spread(Species, mammal_count)

facilities_count
``` 



<div id="w-hint">
**Hint:** You may want to use the **bo** data frame. And eventually the `spread()` function from the **tidyr** package.
</div>

### Turn a wide data frame into a tall tidy data frame

Here's a wide data frame with *Owners* and percent of Belugas compared to Orca whales owned.



```{r display_this}
head(owner_percent)
```


Make this dataframe called **owner_percent** tidy for me (So the mammal types are in one column and the total is in another).

Call the new columns *Species* and *Total*.


```{r t, exercise=TRUE, exercise.lines=4}
# Add to the line below

owner_percent %>%
  
```


```{r t-check}

owner_percent %>%
  gather("Species", "Total", 2:3)

``` 


<div id="t-hint">
**Hint:** You may want to use the `gather()` function.
</div>

### Join Data

Consider these two data sets:

* **owner_percent**

```{r owner_percent}
owner_percent
```

* **owner_total**


```{r owner_total}
owner_total
```

Join these two data sets and call it **owner_total_percent**.

Start with **owner_percent** and have it joined with **owner_total** in a left direction.

```{r j, exercise=TRUE, exercise.lines=4}
# Add to the line below

owner_percent_total <- _____join(________, ________)
  
owner_percent_total
```


```{r j-check}

owner_percent_total <- left_join(owner_percent, owner_total)
  
owner_percent_total

``` 



## Handling strings

Let's load the **stringr** package.

Consider the **bo** data frame again.

```{r string_data}
library(stringr)

bo
```

### Lower case

Let's use the functions in it to manipulate the strings in this data frame.

Turn the names in *Name* to lower case.



```{r low, exercise=TRUE, exercise.lines=4}
bo$Name <- ________(bo$Name)

bo
```


```{r low-check}
bo$Name <- str_to_lower(bo$Name)

bo
``` 


### Title case

Convert the variables in *Sex* to title case.


```{r ti, exercise=TRUE, exercise.lines=4}
bo$Name <- ________(bo$Name)

bo
```


```{r ti-check}
bo$Sex <- str_to_title(bo$Sex)

bo2 <- bo
bo2
``` 


### Combine string

Please create a new variable called *sex_species* that combines *Sex* and *Species* columns with a space in between.

```{r c, exercise=TRUE, exercise.lines=4}
bo3 <- mutate(bo2, 
             _________=_____(_______________________))
bo3
```


```{r c-check}
bo3 <- mutate(bo2,
             sex_species=str_c(Sex, " ", Species))

``` 


### Substitute string

Now, replace all mentions of " Killer Whale;" and " Whale" from the *sex_species* column with a blank.

So "Male Killer Whale; Orca" turns in to "Male Orca" and "Male Beluga Whale" turns into "Male Beluga"


```{r rep, exercise=TRUE, exercise.lines=5}
bo4 <- mutate(bo3, 
             sex_species=________(_______________________)) %>%
      mutate(sex_species=________(_______________________))

bo4$sex_species
```


```{r rep-check}

bo4 <- mutate(bo3,
             sex_species=str_replace(sex_species, " Killer Whale;", "")) %>%
      mutate(sex_species=str_replace(sex_species, " Whale", ""))

bo4$sex_species
``` 


<div id="rep-hint">
**Hint:** You may need to use the str_replace function more than once. Also, pay attention to the order. That will be important.
</div>



### Detect strings

Let's look at the first six rows of the **bo** data frame again.

```{r data_view4}
head(bo)
```


Create a new data frame called **seaworlds** and use filter to keep only the rows that mention "Sea World" in the *Facility* column.


```{r det, exercise=TRUE, exercise.lines=5}
seaworlds <- filter(bo,
                    ______(_____________))

seaworlds
```


```{r det-check}
seaworlds <- filter(bo,
             str_detect(Facility, "Sea World"))

seaworlds
``` 


## Dealing with dates

Consider the original **bo** data frame.

```{r data_view5}
head(bo)
```


### Convert this date

Turn *Captivity.Date* into a date data R can recognize.

We'll use the **lubridate** package.


```{r da, exercise=TRUE, exercise.lines=5}
bo_dates <- bo %>%
  mutate(Captivity.Date=____(_______))

bo_dates
```


```{r da-check}
bo_dates <- bo %>%
  mutate(Captivity.Date=dmy(Captivity.Date))

bo_dates
``` 


### Extract the month

Create a new column called *month_number* and extract the month (as a number) from the *Captivity.Date*.


```{r mo, exercise=TRUE, exercise.lines=5}
bo_dates2 <- bo_dates %>%
  mutate(month_number=____(_______))

bo_dates2
```


```{r mo-check}
bo_dates2 <- bo_dates %>%
  mutate(month_number=month(Captivity.Date))

bo_dates2
``` 



Create a new column called *month_number* and extract the month (as a label) from the *Captivity.Date*.



```{r mo2, exercise=TRUE, exercise.lines=5}
bo_dates3 <- bo_dates2 %>%
  mutate(month_label=____(_______, ______))

bo_dates3
```


```{r mo2-check}
bo_dates3 <- bo_dates2  %>%
  mutate(month_label=month(Captivity.Date, label=T))

bo_dates3
``` 


### Time difference

Consider the modified **bo_dates3** data frame we've created.

```{r last_bo}
bo_dates3
```

Based on available data, can you figure out the average age (in years) of Beluga Whales and Orca Whales (separately) when they join a facility (aka *Facility.Date*)?

Summarize your dataframe (group by *Facility* and *Species*) and create a new column called *average_years*.

```{r lo, exercise=TRUE, exercise.lines=15, warning=F, message=F}
bo_dates3 %>%
```


```{r lo-check}
bo_dates3 %>%
  mutate(Birth.Date=dmy(Birth.Date), 
         Facility.Date=dmy(Facility.Date),
         age=difftime(Facility.Date, Birth.Date),
         age_years=as.numeric(age)/60/60/24/365.25) %>% 
  group_by(Facility, Species) %>% 
  summarize(average_years=mean(age_years, na.rm=T))
  
``` 



<div id="lo-hint">
**Hint:** This one's tough. You'll need to convert *Birth.Date* and *Facility.Date* into dates. Then you'll need to figure out the difference in time between them. And then you'll need to convert that into a number and if it's in seconds, you need to convert that into years. The math for that is seconds/60/60/24/365.25. And then group by how I described. And then when you summarize, be sure when you're finding the average to remove NAs from the calculations. Good luck.
</div>

Alright, we can't really tell the ages because of negative years? It looks like this data is pretty messy. Probably human error.

But great job! You've made it all the way through this chapter!
