---
title: "Chapter 3"
tutorial:
  id: "code.r-journalism/chapter-3"
  version: .8
output:
  learnr::tutorial:
    progressive: true
    theme: lumen
    highlight: espresso
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
knitr::opts_chunk$set(echo = FALSE)
bo <- readRDS("bw.rds")
colnames(bo) <- make.names(colnames(bo))
```

## Transforming and analyzing data

Consider this Marine Mammal Inventory Report data set from the National Oceaninc and Atmospheric Administration.

It has the location of every Beluga and Orca whale in the United States since the '70s.

Currently, the data frame is in the environment and has 494 rows.

This view shows the first six rows.

```{r data_view}
head(bo)
```

Each animal is represented by a *NOAA.ID* (because some do not have names) and might show up multiple times in the data frame.

For example, the Beluga Whale Blanchon is NOA0000719 and has been at three facilities but his latest *Status* shows him as "DEAD".

### Filter

Please create a data frame of Orca whales only from the **bo** data frame.

We'll name the new data frame **orcas**.

```{r filter, exercise=T, exercise.timelimit=60}
# Add to the line below

orcas <- 
  
unique(orcas$Species)
```

### Select

Drop the column/variable *Species* since there's only one variable in it now.

```{r orcas, include=FALSE}
orcas <- filter(bo, Species=="Killer Whale; Orca")
```


```{r select, exercise=T, exercise.timelimit=60}
# Add to the line below

orcas <- 
  

colnames(orcas)
```

### Arrange

Sort the **orcas** data frame based on the *Facility* column in alphabetical order.

```{r arrange, exercise=T, exercise.timelimit=60}
# Add to the line below

orcas <- 
  

head(orcas)
```

### Mutate

Create a new column called *status_num* to fill with a 1 if *Status* is "DEAD" or 0 if alive. Save the transformed data frame back into **orcas**


```{r mutate, exercise=T, exercise.timelimit=60}
# Add to the line below

orcas <- 
  

head(orcas)
```

<div id="mutate-hint">
**Hint:** You may want to use the dplyr `case_when()` function nested in another function.
</div>

### Summarize


```{r orcas2, include=FALSE}
orcas <- mutate(orcas, status_num=case_when(Status=="DEAD" ~ 1, TRUE ~ 0))
```

Count up the number of total Orcas by counting the unique *NOAA.ID*s and how many have died by adding up the *status_num*

```{r summarize, exercise=T, exercise.timelimit=60}
# Add to the line below

alive_dead <- 
  
alive_dead
```


<div id="summarize-hint">
**Hint:** You may want to use the dplyr `n_distinct()` function nested in another function.
</div>

### Percent


What's the percent of dead Orcas?

```{r orcas3, include=FALSE}
alive_dead <- mutate(orcas, status_num=case_when(Status=="DEAD" ~ 1, TRUE ~ 0)) %>% 
  summarize(total=n_distinct(NOAA.ID), died=sum(status_num))
```

Now figure how the percent of Orcas still alive.

```{r mutate2, exercise=T, exercise.timelimit=60}
# Add to the line below

percent_alive <-

percent_alive
```

### Group By

Which facilities have had the most Orcas and Belugas combined? 

List them from highest count to lowest.

```{r facilities_count, exercise=T, exercise.timelimit=60}
# Add to the line below

facilities_count <-

facilities_count
```


<div id="facilities_count-hint">
**Hint:** You may want to use the **bo** data frame. Double check the variables available above.
</div>

## Tidying and joining data

Consider the **bo** data set again.

```{r data_view2}
head(bo)
```

### Turn a tall data frame wide

Summarize the data and count the *total* number of Orcas and Belugas per facility from the *Species* variable.

Then turn the tall data into a wide data frame that has these columns: Facility, Belugas, Orcas.

```{r wide, exercise=T, exercise.timelimit=60}
# Add to the line below

facilities_count <- bo %>%
   group_by(Facility, ________) %>%
   summarize(_______=___) %>%
   ______(_______, _____)

facilities_count
```

<div id="wide-hint">
**Hint:** You may want to use the **bo** data frame. And eventually the `spread()` function from the **tidyr** package.
</div>

### Turn a wide data frame into a tall tidy data frame

Here's a wide data frame with *Owners* and percent of Belugas compared to Orca whales owned.


```{r clean_up, include=F}
owner_percent <- bo %>%
   group_by(Owner, Species) %>%
   summarize(total=n()) %>%
   mutate(percent=round(total/sum(total, na.rm=T)*100,2)) %>%
   select(-total) %>%
   spread(Species, percent)
```


```{r display_this}
head(owner_percent)
```


Make this dataframe called **owner_percent** tidy for me.

```{r tall, exercise=T, exercise.timelimit=60}
# Add to the line below

owner_percent %>%
  
  
```

<div id="tall-hint">
**Hint:** You may want to use the `gather()` function.
</div>

### Join Data

Consider these two data sets:

* **owner_percent**

```{r join_left, include=F}
owner_percent <- bo %>%
   group_by(Owner, Species) %>%
   summarize(total=n()) %>%
   mutate(percent=round(total/sum(total, na.rm=T)*100,2)) %>%
   select(-total)
```

```{r owner_percent}
owner_percent
```

* **owner_total**

```{r join_right, include=F}
owner_total <- bo %>%
   group_by(Owner, Species) %>%
   summarize(total=n())
```

```{r owner_total}
owner_total
```

Join these two data sets and call it **owner_total_percent**.

```{r join, exercise=T, exercise.timelimit=60}
# Add to the line below

owner_total_percent <- _________(________, ________)
  
owner_total_percent
```


## Handling strings

Let's load the **stringr** package.

Consider the **bo** data frame again.

```{r string_data}
library(stringr)

bo
```

### Lower case

Let's use the functions in it to manipulate the strings in this data frame.

Turn the names in *Name* to lower case.

```{r lowercase, exercise=T, exercise.timelimit=60}
bo$Name <- ________(bo$Name)

bo
```

```{r name_lower, include=F}
bo$Name <- str_to_lower(bo$Name)
```

### Title case

Convert the variables in *Sex* to title case.

```{r titlecase, exercise=T, exercise.timelimit=60}
bo$Sex <- ________(bo$Sex)

bo
```


```{r name_title, include=F}
bo$Name <- str_to_title(bo$Name)
```

### Combine string

Please create a new variable called *sex_species* that combines *Sex* and *Species* columns with a space in between.

```{r combine, exercise=T, exercise.timelimit=60}
bo <- mutate(bo, 
             _________=_____(_______________________))
bo
```


```{r combined, include=F}
bo <- mutate(bo,
             sex_species=str_c(Sex, " ", Species))

```
### Substitute string

Now, replace all mentions of " Killer Whale;" and " Whale" from the *sex_species* column with a blank.

So "Male Killer Whale; Orca" turns in to "Male Orca" and "Male Beluga Whale" turns into "Male Beluga"



```{r replace, exercise=T, exercise.timelimit=60}
bo <- mutate(bo, 
             sex_species=________(_______________________)) %>%
      mutate(sex_species=________(_______________________))

bo$sex_species
```



<div id="replace-hint">
**Hint:** You may need to use the str_xxxxxxxx function more than once. Also, pay attention to the order. That will be important.
</div>



```{r replaced, include=F}
bo <- mutate(bo,
             sex_species=str_replace(sex_species, " Killer Whale;", "")) %>%
      mutate(sex_species=str_replace(sex_species, " Whale", ""))
```

### Detect strings

Create a new data frame called **seaworlds** and filter out rows that mention "Sea World" in the *Facility* column.

```{r detect_string, exercise=T, exercise.timelimit=60}
seaworlds <- filter(bo,
                    ______(_____________))

seaworlds
```


```{r seaworlds, de=F}
seaworlds <- filter(bo,
             str_detect(Facility, "Sea World"))

```

## Dealing with dates

Consider the modified **bo** data frame we've created.

### Convert this date

Turn *Captivity.Date* into a date data R can recognize.

We'll use the **lubridate** package.

```{r lubridate, include=F}
library(lubridate)
```


```{r date, exercise=T, exercise.timelimit=60}
bo <- bo %>%
  mutate(Captivity.Date=____(_______))

bo
```


```{r dated, include=F}
bo <- bo %>%
  mutate(Captivity.Date=dmy(Captivity.Date))

```

### Extract the month

Create a new column called *month_number* and extract the month (as a number) from the *Captivity.Date*.

```{r date_month, exercise=T, exercise.timelimit=60}
bo <- bo %>%
  mutate(month=____(_______))

bo
```


```{r dated_month, include=F}
bo <- bo %>%
  mutate(month_number=month(Captivity.Date))

```

Create a new column called *month_number* and extract the month (as a number) from the *Captivity.Date*.

```{r date_month2, exercise=T, exercise.timelimit=60}
bo <- bo %>%
  mutate(month_label=____(_______))

bo
```


```{r dated_month2, include=F}
bo <- bo %>%
  mutate(month_label=month(Captivity.Date, label=T))

```

### Time difference

Consider the modified **bo** data frame we've created.

```{r last_bo}
bo
```

Based on available data, can you figure out the average age (in years) of Beluga Whales and Orca Whales in when they join a facility?

```{r last_analysis, exercise=T, exercise.timelimit=180}
bo %>%
  
  
  
  
  
  
  
  
  
  
```