---
title: "Chapter 4"
tutorial:
  id: "code.r-journalism/chapter-4"
  version: .85
output:
  learnr::tutorial:
    progressive: true
    theme: lumen
    highlight: espresso
    include:
      before_body: _navbar.html
runtime: shiny_prerendered
---


```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(forcats)
#install.packages("wesanderson")
library(wesanderson)
library(ggrepel)

knitr::opts_chunk$set(echo = FALSE)
movies <- readRDS("movies.rds") %>%
  filter(country=="USA") %>% 
  mutate(genre1=gsub("\\|.*", "", genres)) %>% 
  filter(content_rating %in% c("G", "PG", "PG-13", "R")) %>% 
  filter(title_year>1982)



nospace <- function(a) {
  a <- gsub(" ","",a) 
  a <- gsub("\\\n", "", a)
  gsub("\\`", "", a)
}


is_bad_code <- function(user_code, check_code, envir_result) {

  is_null <- is.null(eval(parse(text = user_code)))
  is_gg <- any(grepl("gg", class(eval(parse(text = user_code)))))
  e_rows <- nrow(eval(parse(text = user_code)))
  r_rows <- nrow(eval(parse(text = check_code), envir = envir_result))
  e_cols <- ncol(eval(parse(text = user_code)))
  r_cols <- ncol(eval(parse(text = check_code), envir = envir_result))
  e_type <- typeof(eval(parse(text = user_code)))
  r_type <- typeof(eval(parse(text = check_code), envir=envir_result))
  e_len <- length(eval(parse(text = user_code)))
  r_len <- length(eval(parse(text = check_code)))
  if (is_null!=T & is_gg!=T){
    if (e_len!=0 & r_len!=0) {
      if (e_type==r_type) {
        if (e_type!="character" & e_type!="double" & e_type!="integer" & e_type!="logical") {
          if (e_rows==r_rows && e_cols==r_cols) {
            eval(parse(text = user_code)) != eval(parse(text = check_code), envir = envir_result)
          } else {
            TRUE
          }
        } else { eval(parse(text = user_code)) != eval(parse(text = check_code), envir = envir_result) }
      } else {
        TRUE
      }
    } else { TRUE }
  } else {nospace(user_code) != nospace(check_code)}
}


checker <- function(label, user_code, check_code, envir_result, evaluate_result, ...) {
  if (is_bad_code(user_code, check_code, envir_result)) {
    return(list(message = "Code wasn't right!", correct = FALSE, location = "append"))
  } else {
    return(list(message = "Great Job!", correct = TRUE, location = "append"))
  }
}  

tutorial_options(exercise.timelimit = 30, exercise.checker = checker)

```

## ggplot2

Consider this data set from IMDB on 5,000 movies we have in the environment called **movies**.


```{r display}
movies
```

This is what `glimpse(movies)` outputs.

```{r glimpse}
glimpse(movies)
```


### Make a scatterplot

Let's start easy with a simple scatter plot comparing box office gross to the budget.


```{r v2, exercise=TRUE}
ggplot(______) + geom_________(aes(x=____,y=_____))

```


```{r v2-check}
ggplot(movies) + geom_point(aes(x=gross,y=budget))
``` 


### Change the color

Let's change the color of the circles to blue.

```{r s1, exercise=TRUE, warning=F, message=F}
ggplot(______) + geom_________(aes(x=_____,y=_____), ________)

```


```{r s1-check}
ggplot(movies) + geom_point(aes(x=gross,y=budget), color="blue")
``` 



### Make a scatterplot with color groups

Add a color factor based on `content_rating`.


```{r sc1, exercise=TRUE, warning=F, message=F}
ggplot(______) + geom_________(aes(x=_____, y=______, _______=_____))

```


```{r sc1-check}
ggplot(movies) + geom_point(aes(x=gross,y=budget, color=content_rating))
``` 


Did you notice the placement of the second-to-last parenthesis for color this time?
  
### Bar plot
  
Make a bar plot chart that counts up the number of titles per year (*title_year*).


```{r bp1, exercise=TRUE, warning=F, message=F}
ggplot(______,
             aes(x=_________)) +
  geom__________()
```


```{r bp1-check}
ggplot(movies,
             aes(x=title_year)) +
  geom_bar()
``` 


### Stacked bar plot ver. 1

Add *content_rating* as a grouping per year counting up the movies to create a stacked bar chart.


```{r bp2, exercise=TRUE, warning=F, message=F}
ggplot(______,
             aes(x=_________,_________)) +
  geom__________()
```


```{r bp2-check}
ggplot(movies,
             aes(x=title_year, fill=content_rating)) +
  geom_bar()
``` 


<div id="bp2-hint">
  **Hint:** You may want to use the `fill` argument in the `aes()`.
</div>
  
### Stacked bar plot ver. 2
  
Great, now split up the bars so they're not stacked but next to each other.

And we'll focus on movies created after 2001 (*title_year* is the variable).


```{r bp3, exercise=TRUE, warning=F, message=F}
movies %>% 
  filter(___________) %>%
  ggplot(aes(x=_________,fill=________)) +
  geom__________(________________)
```


```{r bp3-check}
movies %>% 
  filter(title_year>2001) %>%
  ggplot(aes(x=title_year,fill=content_rating)) +
  geom_bar(position="dodge")
``` 


<div id="bp3-hint">
  **Hint:** You may want to use the `position` argument in the `geom_bar()` function.
</div>
  
### Stacked bar plot ver. 3
  
Alright, let's make a percent stacked chart this time:


```{r bp4, exercise=TRUE, warning=F, message=F}
movies %>% 
  filter(___________) %>%
  ggplot(aes(x=_________,fill=_________)) +
  geom__________(position=________)
```


```{r bp4-check}
movies %>% 
  filter(title_year>2001) %>%
  ggplot(aes(x=title_year,fill=content_rating)) +
  geom_bar(position="fill")
``` 


## Customizing charts

### Another bar chart

Consider this data set from IMDB on 5,000 movies we have in the environment called **movies**.


```{r display2}
movies
```

This is what `glimpse(movies)` outputs.

```{r glimpse2}
glimpse(movies)
```

Let's look at box office results for all the movies that James Cameron has created (variable is *director_name*).

```{r jc, exercise=TRUE, warning=F, message=F}
movies %>% 
  filter(__________) %>% 
  ggplot(aes(x=_________,y=___________)) +
  geom_bar(___________)
```


```{r jc-check}
movies %>% 
  filter(director_name=="James Cameron") %>%
  ggplot(aes(x=movie_title,y=gross)) +
  geom_bar(stat="identity")
``` 


<div id="jc-hint">
  **Hint:** You may want to pass the argument `stat=` to the `geom_bar()` function. What do you fill with stat? You'll need to check your [notes](https://learn.r-journalism.com/en/visualizing/charts_with_ggplot/ggplot2/).
</div>

### Flip that chart

Transpose that chart so that the movies are on the y axis instead of the x axis (without swapping the coords from the code above).

```{r j_c2, exercise=TRUE, warning=F, message=F}
movies %>% 
  filter(__________) %>% 
  ggplot(aes(x=movie_title,y=gross)) +
  geom_bar(___________) +
  __________()
```


```{r j_c2-check}
movies %>% 
  filter(director_name=="James Cameron") %>%
  ggplot(aes(x=movie_title,y=gross)) +
  geom_bar(stat="identity") +
  coord_flip()
``` 





<div id="j_c2-hint">
**Hint:** What's that line you need to add? You're **flip**ping the **coord**inates.
</div>

### Reorder the labels in the chart

Let's recreate the chart so that the movies listed are in order of release date (variable is *title_year*).

Remember, you'll need to use the library **forcats** and the function `fct_reorder()`.


```{r j_c3, exercise=TRUE, warning=F, message=F}
movies %>% 
  filter(__________) %>%
  ggplot(aes(x=fct_reorder(_________,____(_______)), y=___________)) +
  geom_bar(___________) +
  __________()
```


```{r j_c3-check}
movies %>% 
  filter(director_name=="James Cameron") %>%
  ggplot(aes(x=fct_reorder(movie_title,desc(title_year)), y=gross)) +
  geom_bar(stat="identity") +
  coord_flip()
``` 





<div id="j_c3-hint">
**Hint:** I've already given you the `fct_reorder()` function! Just [look up](http://learn.r-journalism.com/en/visualizing/customizing_charts/customizing-exporting-ggplot2/) how it works so you use it right.
</div>
  
### Style it

Let's fix the x-axis and y-axis labels ("Movie" and "Box Office Gross"), as well as add a title to the chart ("How James Cameron movies performed at the box office") and caption for where the data came from ("Source: IMDB.com")



```{r j_c4, exercise=TRUE, warning=F, message=F}
movies %>% 
  filter(____________) %>%
  ggplot(aes(x=fct_reorder(__________,desc(_______)), y=_____)) +
  geom_bar(stat="identity") +
  coord_flip() +
  ____(x=___________,
       y=___________,
       title=___________,
       sub____=___________) +
  theme_minimal()
```


```{r j_c4-check}
movies %>% 
  filter(director_name=="James Cameron") %>%
  ggplot(aes(x=fct_reorder(movie_title,desc(title_year)), y=gross)) +
  geom_bar(stat="identity") +
  coord_flip() +
  labs(x="Movie",
       y="Box Office Gross",
       title="How James Cameron movies performed at the box office",
       subtitle="Source: IMDB.com") +
  theme_minimal()
``` 



<div id="j_c4-hint">
**Hint:** You might want to use the `labs()` function. But you'll have to [look up](http://learn.r-journalism.com/en/visualizing/customizing_charts/customizing-exporting-ggplot2/) the rest.
</div>
  
  
### Color palettes
  
Let's look at Wes Anderson's box office performance using the code as above, but with Wes Anderson substituted out for James Cameron.

And for fun, we'll use the **wesanderson** color palette from his first movie, Bottle Rocket (*BottleRocket1*).

Because you're changing the color of the bars based on the *movie_title* variable, you need to add that to the `aes()`.

Keep the movie title similar to the previous chart but with Wes Anderson instead of James Cameron.

```
library(wesanderson)
```


```{r j_c5, exercise=TRUE, warning=F, message=F}
movies %>% 
  filter(director=="Wes Anderson") %>%
  ggplot(aes(x=fct_reorder(_________,desc(_______)), y=___________, ____=_____________)) +
  geom_bar(___________) +
  __________() +
  ____(____=___________,
           ____=___________,
           ______=__________________,
           ______=___________________) +
  theme_minimal() +
  ______________(values=wes_palette("_________"), guide=F)
```


```{r j_c5-check}
movies %>% 
  filter(director_name=="Wes Anderson") %>%
  ggplot(aes(x=fct_reorder(movie_title,desc(title_year)), y=gross, fill=movie_title)) +
  geom_bar(stat="identity") +
  coord_flip() +
  labs(x="Movie",
       y="Box Office Gross",
       title="How Wes Anderson movies performed at the box office",
       subtitle="Source: IMDB.com") +
  theme_minimal() +
  scale_fill_manual(values=wes_palette("BottleRocket1"), guide=F)
``` 



<div id="j_c5-hint">
  **Hint:** In the last line, you're using a scale function that lets you change the fill colors manually. Here's the [documentation](https://github.com/karthik/wesanderson) on the Wes Anderson palette
</div>
  
### Labeling
  
Let's filter these movies to those directed by "Kathryn Bigelow", "Martin Scorsese", and "Steven Spielberg" and then make a scatter plot comparing box office to budget. 

The color should represent the different director.

Then add a label for the movie using the [**ggrepel**](https://cran.r-project.org/web/packages/ggrepel/vignettes/ggrepel.html) package.

```
library(ggrepel)
```


```{r ggr, exercise=TRUE, warning=F, message=F, fig.width=7, fig.height=7}
movies %>% 
  filter(director_name ____ c("Kathryn Bigelow", "Martin Scorsese", "Steven Spielberg")) %>% 
  ggplot(aes(x=gross, y=budget, color=______________, ______=___________)) +
  geom_point() +
  geom_____________()

```


```{r ggr-check}
movies %>% 
  filter(director_name %in% c("Kathryn Bigelow", "Martin Scorsese", "Steven Spielberg")) %>% 
  ggplot(aes(x=gross, y=budget, color=director_name, label=movie_title)) +
  geom_point() +
  geom_text_repel()
``` 




<div id="ggr-hint">
**Hint:** The filter code that lets you filter by group is `%in%`. The color is based on the director and the label is based on the movie name. Can you find those variable names in the **movies** data frame? And what's the function for the last line? If you forgot, refer to [the lesson](http://learn.r-journalism.com/en/visualizing/more_charts/fancy-charts/).
</div>
  
  
### Small multiples
  
Nice! 

But crowded.

Let's break them out individually by director so it's easier to see the patterns.


```{r ggrf, exercise=TRUE, warning=F, message=F, fig.width=6, fig.height=6}
movies %>% 
  filter(director_name ____ c("Kathryn Bigelow", "Martin Scorsese", "Steven Spielberg")) %>% 
  ggplot(aes(x=gross, y=budget, color=______________, ______=___________)) +
  geom_point() +
  geom_____________(size=2)
_____________(~____________, nrow=2) +
  theme(legend.position="none")
```


```{r ggrf-check}
movies %>% 
  filter(director_name %in% c("Kathryn Bigelow", "Martin Scorsese", "Steven Spielberg")) %>% 
  ggplot(aes(x=gross, y=budget, color=director_name, label=movie_title)) +
  geom_point() +
  geom_text_repel(size=2) +
  facet_wrap(~director_name, nrow=2) +
  theme(legend.position="none")
``` 

<div id="ggrf_facet-hint">
  **Hint:** What's the term for creating facets of the chart? It's facet_*something*. There's [more than one](http://learn.r-journalism.com/en/visualizing/charts_with_ggplot/ggplot2/), but we want the one that we can specify the number of rows in the output (with `nrow`).
</div>


