---
title: "Chapter 5"
tutorial:
  id: "code.r-journalism/chapter-5"
  version: .8
output:
  learnr::tutorial:
    progressive: true
    theme: lumen
    highlight: espresso
    include:
      before_body: _navbar.html
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(sf)
library(leaflet)
library(viridis)

knitr::opts_chunk$set(echo = FALSE)

map_filepath <- "Boston_Neighborhoods/Boston_Neighborhoods.shp"

boston <- st_read(map_filepath)

homicides <- read_csv("homicide-data.csv")
homicides <- homicides %>% 
  filter(city=="Boston") %>% 
  filter(!is.na(lon))

race_ethnicity <- read_csv("race-and-or-ethnicity.csv")
colnames(race_ethnicity) <- make.names(colnames(race_ethnicity))
race_ethnicity$Percent.of.Population <- as.numeric(gsub("\\%", "", race_ethnicity$Percent.of.Population))
race_ethnicity2010 <- filter(race_ethnicity, Decade==2010)

boston_population <- race_ethnicity2010 %>% group_by(Neighborhood) %>% 
  summarize(Population=sum(Number.of.People))

boston_race_map <- left_join(race_ethnicity2010, boston, by=c("Neighborhood"="Name"))

boston_pop_map <- left_join(boston_population, boston, by=c("Neighborhood"="Name"))


boston_black_perc <- readRDS("boston_black.rds")

black_boston <- boston_black_perc %>% 
  rename(black_percent=black_residents) %>% 
  filter(black_percent > 50)

murders_spatial <- homicides %>% 
  st_as_sf(coords=c("lon", "lat"), crs = "+proj=longlat") %>% 
  st_transform(crs=st_crs(boston))

points_in <- st_join(boston, murders_spatial, left=T)

disp <- colorFactor(c("#ff0000", "#13ED3F"), domain=c("Closed by arrest", "Open/No arrest"))

nospace <- function(a) {
  a <- gsub(" ","",a) 
  a <- gsub("\\\n", "", a)
  gsub("\\`", "", a)
}


is_bad_code <- function(user_code, check_code, envir_result) {

  is_null <- is.null(eval(parse(text = user_code)))
  is_gg <- any(grepl("gg", class(eval(parse(text = user_code)))))
  e_class <- class(eval(parse(text = user_code)))[1]
  r_class <- class((eval(parse(text = check_code), envir = envir_result)))[1]
  e_rows <- nrow(eval(parse(text = user_code)))
  r_rows <- nrow(eval(parse(text = check_code), envir = envir_result))
  e_cols <- ncol(eval(parse(text = user_code)))
  r_cols <- ncol(eval(parse(text = check_code), envir = envir_result))
  e_type <- typeof(eval(parse(text = user_code)))
  r_type <- typeof(eval(parse(text = check_code), envir=envir_result))
  e_len <- length(eval(parse(text = user_code)))
  r_len <- length(eval(parse(text = check_code)))
  if (is_null!=T & is_gg!=T & r_class!="sf" & r_class!="tbl_df" & r_class!="leaflet"){
    if (e_len!=0 & r_len!=0) {
      if (e_type==r_type) {
        if (e_type!="character" & e_type!="double" & e_type!="integer" & e_type!="logical") {
          if (e_rows==r_rows && e_cols==r_cols) {
            eval(parse(text = user_code)) != eval(parse(text = check_code), envir = envir_result)
          } else {
            TRUE
          }
        } else { eval(parse(text = user_code)) != eval(parse(text = check_code), envir = envir_result) }
      } else {
        TRUE
      }
    } else { TRUE }
  } else {nospace(user_code) != nospace(check_code)}
}


checker <- function(label, user_code, check_code, envir_result, evaluate_result, ...) {
  if (is_bad_code(user_code, check_code, envir_result)) {
    return(list(message = "Code wasn't right!", correct = FALSE, location = "append"))
  } else {
    return(list(message = "Great Job!", correct = TRUE, location = "append"))
  }
}  

tutorial_options(exercise.timelimit = 30, exercise.checker = checker)

```

## Shape files

### Load this shapefile

The libraries you need for working with maps are already loaded.

```
library(ggplot2)
library(sf)
```

Let's load in a shapefile we have on the computer locally.

It's in the folder *Boston_Neighborhoods* and the shapefile is also called *Boston_Neighborhoods*. Load it in.

Call the object *boston*.

```
library(sf)
library(ggplot2)
library(dplyr)

map_filepath <- "Boston_Neighborhoods/Boston_Neighborhoods.shp"

boston <- st_read(map_filepath)
```

### Glimpse

Let's take a look at the data behind the shapefile.

```{r glimpse, exercise=TRUE}
_____(boston)
```

```{r glimpse-check}
glimpse(boston)
``` 

### Plot the shapefile

```{r plot, exercise=TRUE}
_____(_______) +
  _________()
```

```{r plot-check}
ggplot(boston) +
  geom_sf()
``` 


<div id="plot-hint">
The ggplot2 and sf function you need to map out a shapefile is geom_sf().
</div>

Good job. Not much to display, yet.

So let's bring in some data to join it with.

### Load data

The data is historical population and race data by neighborhood from the City of Boston. 

It's loaded in your environment as **boston_population**.

Take a glimpse.

```{r glimpse2, exercise=TRUE}
_____(boston_population))
```

```{r glimpse2-check}
glimpse(boston_population)
```

### Join the shapefile to the data

Combine the shape file **boston** to the data frame **boston_population**.

Use the function that joins from the left from *dplyr*.

```{r join, exercise=TRUE, warning=F}
boston_pop_map <- _____(______, _______, by=c(_______=_______))

boston_pop_map
```

```{r join-check}
boston_pop_map <- left_join(boston_population, boston, by=c("Neighborhood"="Name"))

boston_pop_map
```

<div id="join-hint">
The dplyr verb you need to use to join to the left is left_join() and don't forget that variable names in by need to be in quotes.
</div>

### Plot the data on the shapefile

Make a choropleth with colors based on *Population*.

```{r plot2, exercise=TRUE, warning=F}
ggplot(boston_pop_map) +
  geom_sf(aes(fill=________))
``` 


```{r plot2-check}
ggplot(boston_pop_map) +
  geom_sf(aes(fill=Population))
``` 

### Style it

First, let's fix the legend so the gradient goes in the other direction.

And also add a title "Boston neighborhood population" and source in the caption.

```{r plot3, exercise=TRUE, warning=F}
ggplot(boston_pop_map) +
  geom_sf(aes(fill=Population)) +
  scale_fill__________(_________=1, _______="Population") +
  _____(_________=________________, _______="Source: City of Boston")
``` 


```{r plot3-check}
ggplot(boston_pop_map) +
  geom_sf(aes(fill=Population))  +
  scale_fill_distiller(direction=1, name="Population") +
  labs(title="Boston neighborhood population", caption="Source: US Census")
``` 


<div id="plot3-hint">
scale_fill_distiller()
</div>

Also, let's get rid of the axis marks and the background gray grid.

```{r plot3a, exercise=TRUE, warning=F}
ggplot(boston_pop_map) +
  geom_sf(aes(fill=Population)) +
  scale_fill__________(_________=1, _______="Population") +
  _____(_________=________________, _______="Source: City of Boston") +
  theme________() +
  theme(_________________ = element_line(colour = ____________)) +

``` 


```{r plot3a-check}
ggplot(boston_pop_map) +
  geom_sf(aes(fill=Population))  +
  scale_fill_distiller(direction=1, name="Population") + 
  labs(title="Boston neighborhood population", caption="Source: US Census") +
  theme_void() +
  theme(panel.grid.major = element_line(colour = 'transparent'))
``` 



<div id="plot3a-hint">
These style tweaks are in [Chapter 5](http://learn.r-journalism.com/en/mapping/static_maps/static-maps/).
</div>

## Small multiples

### More data

Let's map historical population and race data by neighborhood from the City of Boston. 

It's loaded in your environment as **race_ethnicity2010**.

Take a glimpse.

```{r glimpse4, exercise=TRUE}
_____(race_ethnicity2010)
```

```{r glimpse4-check}
glimpse(race_ethnicity2010)
```

### Join the shapefile to the data

Combine the shape file **boston** to the data frame **race_ethnicity2010**.

Use the function that joins from the left from *dplyr*.

```{r join2, exercise=TRUE, warning=F}
boston_race_map <- _____(race_ethnicity2010, _______, by=c(_______=_______))

boston_race_map
```


```{r join2-check}
boston_race_map <- left_join(race_ethnicity2010, boston, by=c("Neighborhood"="Name"))

boston_race_map
```



### Plot the data on the shapefile

Map the data, setting the color based on *Percent.of.Population*. 

Also, facet wrap out by *Race.and.or.Ethnicity* this time.

**This will take a few moments to render**.

```{r plot4, exercise=TRUE, warning=F}
ggplot(boston_race_map) +
  geom_sf(aes(fill=________)) +
  __________(~___________, ncol=2)
``` 


```{r plot4-check}
ggplot(boston_race_map) +
  geom_sf(aes(fill=Percent.of.Population)) +
  facet_wrap(~Race.and.or.Ethnicity, ncol=2)
``` 

### Style with title and color fix

First, let's fix the legend so the gradient goes in the other direction.

And also add a title "Boston neighborhood population by race" and source in the caption.

**This will take a few moments to render**.

```{r plot5, exercise=TRUE, warning=F}
ggplot(boston_race_map) +
  geom_sf(aes(fill=Percent.of.Population)) +
    __________(~___________, ncol=2) +
  scale_fill__________(_________=1, _______="Percent of Population") +
  _____(_________=________________, _______="Source: City of Boston")
``` 


```{r plot5-check}
ggplot(boston_race_map) +
  geom_sf(aes(fill=Percent.of.Population)) +
  facet_wrap(~Race.and.or.Ethnicity, ncol=2) +
  scale_fill_distiller(direction=1, name="Percent of Population") +
  labs(title="Boston neighborhood population by race", caption="Source: City of Boston")
``` 

### Style with grid and axis fixes

Also, let's get rid of the axis marks and the background gray grid.

**This will take a few moments to render**.

```{r plot6, exercise=TRUE, warning=F}
ggplot(boston_race_map) +
  geom_sf(aes(fill=Percent.of.Population)) +
  scale_fill__________(_________=1, _______="Percent of Population") +
  _____(_________=________________, _______="Source: City of Boston") +
  theme________() +
  theme(_________________ = element_line(colour = ____________)) +

``` 


```{r plot6-check}
ggplot(boston_race_map) +
  geom_sf(aes(fill=Percent.of.Population))  +
  scale_fill_distiller(direction=1, name="Percent of Population") + 
  facet_wrap(~Race.and.or.Ethnicity, ncol=2) +
  labs(title="Boston neighborhood population by race", caption="Source: City of Boston") +
  theme_void() +
  theme(panel.grid.major = element_line(colour = 'transparent'))
``` 

## Mapping murders

This data comes from The Washington Post's [Murder with Impunity](https://www.washingtonpost.com/graphics/2018/investigations/where-murders-go-unsolved/) series.

We'll start out with a slice of Boston's data.

Take a glimpse at the structure of the data, which is in the environment as **homicides**.

```{r glimpse5, exercise=TRUE}
_____(homicides)
```

```{r glimpse5-check}
glimpse(homicides)
```

### Plot the points on the map

Let's use the leaflet package to map out all the locations of the murders in Boston.

We'll use the **boston** shape file as the base. Then add the locations based on the latitude and longitude columns in the **homicides** data frame.

```
library(leaflet)
```

Fill in the blanks to 1) add the shapefile of Boston neighborhoods. Remember, the shapefile is in the global environment as **boston**. Then, 2) Add circles for every longitude and latitude in the **homicides** data frame and have the popup be the *disposition* variable.


```{r leaflet, exercise=TRUE, warning=F}
leaflet(________) %>% addProviderTiles("CartoDB.Positron") %>% 
  setView(-71.087010, 42.316240, zoom = 11) %>% 
    __________( 
              fillColor = "transparent", 
              fillOpacity = 0.9, 
              weight = 1, 
              smoothFactor = 0.2,
              color="black") %>% 
    add_______(~________$____, ~_________$__, popup=______$______, weight = 3, 
               radius=40, color="#ff0000", stroke = TRUE, fillOpacity = 0.8) 
```

```{r leaflet-check}
leaflet(boston) %>% addProviderTiles("CartoDB.Positron") %>% 
  setView(-71.087010, 42.316240, zoom = 11) %>% 
    addPolygons( 
              fillColor = "transparent", 
              fillOpacity = 0.9, 
              weight = 1, 
              smoothFactor = 0.2,
              color="black") %>% 
    addCircles(~homicides$lon, ~homicides$lat, popup=homicides$disposition, weight = 3, radius=40, color="#ff0000", stroke = TRUE, fillOpacity = 0.8) 
```

### Points in a polygon prep

We want to associate each dot with a corresponding neighborhood.

Prepare points for join by making points fit the projection of the  shapefile it will be joined with.


```{r projection, exercise=TRUE, warning=F}
murders_spatial <- homicides %>% 
  ________(coords=c("lon", "lat"), crs = "+proj=longlat") %>% 
  __________(crs=_____(boston)) 

str(murders_spatial)
```


```{r projection-check}
murders_spatial <- homicides %>% 
  st_as_sf(coords=c("lon", "lat"), crs = "+proj=longlat") %>% 
  st_transform(crs=st_crs(boston))

str(murders_spatial)
```


<div id="pointsin-hint">
You'll be using three different st_xxxxx functions. If you need reference, we've [used it before](
http://learn.r-journalism.com/en/mapping/case_study/traffic-stops/).
</div>

### Points in a polygon join

Use the special function that sees where the geometries we’ve set in **murders_spatial** fall into which polygon in **boston**.

```{r pointsin, exercise=TRUE, warning=F}
points_in <- __________(boston, _________, left=T)

points_in
```

```{r pointsin-check}
points_in <- st_join(boston, murders_spatial, left=T)

points_in
```

We'll bring in the viridis library for colors.

```
library(viridis)
```

And now we can make a choropleth map with this new transformed data frame.

Start with the **points_in** data frame and figure out the number of murders per neighborhood (**Name** is the variable name). You'll need to create a new variable called *murders*.



```{r choro1, exercise=TRUE, warning=F}
points_in %>% 
  group_by(_____) %>%
  __________(_________=___) %>% 
________() +
  ________(aes(fill = ________), color=NA) +
  coord_sf(datum=NA) +
  labs(title = "Total murders in Boston by neighborhood",
       subtitle = "Between 2008 and 2018",
       caption = "Source: Washington Post",
       fill = "Murders") +
  scale_fill_viridis(option="magma", direction=-1)
```

```{r choro1-check}
points_in %>% 
  group_by(Name) %>%
  summarize(murders=n()) %>% 
ggplot() +
  geom_sf(aes(fill = murders), color=NA) +
  coord_sf(datum=NA) +
  labs(title = "Total murders in Boston by neighborhood",
       subtitle = "Between 2008 and 2018",
       caption = "Source: Washington Post",
       fill = "Murders") +
  scale_fill_viridis(option="magma", direction=-1)

```

Nice!

Now, let's figure out percent of unsolved murders by neighborhood and map that out.

You'll have to add another variable (*disposition*) to group by and after summarizing by creating a new variable (*murders*), you'll then create another variable *percent* that figures out the percent.

And then filter disposition for `"Open/No arrest"`. 

Then plot it with the function from ggplot2().

```{r choro2, exercise=TRUE, warning=F}
points_in %>% 
  ______(____, ________) %>%
  ________(________=___) %>% 
  ________(______=______/___(_________)*100) %>% 
  filter(____________=="Open/No arrest") %>% 
ggplot() +
  _______(aes(fill = _______), color=NA) +
  coord_sf(datum=NA) +
  labs(title = "Rate of unsolved murders in Boston by neighborhood",
       subtitle = "Between 2008 and 2018",
       caption = "Source: Washington Post",
       fill = "Percent unsolved") +
  scale_fill_viridis(option="magma", direction=-1)

```


```{r choro2-check}
points_in %>% 
  group_by(Name, disposition) %>%
  summarize(murders=n()) %>% 
  mutate(percent=murders/sum(murders)*100) %>% 
  filter(disposition=="Open/No arrest") %>% 
ggplot() +
  geom_sf(aes(fill = percent), color=NA) +
  coord_sf(datum=NA) +
  labs(title = "Rate of unsolved murders in Boston by neighborhood",
       subtitle = "Between 2008 and 2018",
       caption = "Source: Washington Post",
       fill = "Percent unsolved") +
  scale_fill_viridis(option="magma", direction=-1)

```

<div id="choro2-hint">
Remember, to find the percent of something, it's `something/sum(something)*100`.
</div>

Let's see if we can do the analysis above but for each ethnicity listed in Boston.

Quickly take a look at the **points_in** data frame to look up what variable you want to add.

```{r glimpsed}
glimpse(points_in)
```

Now add that variable into the group_by and make a small multiples viz of the map based on that new variable.

```{r choro3, exercise=TRUE, warning=F}
points_in %>% 
  group_by(Name, __________, disposition) %>%
  summarize(murders=n()) %>% 
  mutate(percent=murders/sum(murders)*100) %>% 
  filter(disposition=="Open/No arrest") %>% 
ggplot() +
  geom_sf(aes(fill = percent), color=NA) +
  coord_sf(datum=NA) +
  labs(title = "Rate of solved murders in Boston by neighborhood",
       subtitle = "Between 2008 and 2018",
       caption = "Source: Washington Post",
       fill = "Percent unsolved") +
  scale_fill_viridis(option="magma", direction=-1) +
  __________(~_________)
```



```{r choro3-check}
points_in %>% 
  group_by(Name, victim_race, disposition) %>%
  summarize(murders=n()) %>% 
  mutate(percent=murders/sum(murders)*100) %>% 
  filter(disposition=="Open/No arrest") %>% 
ggplot() +
  geom_sf(aes(fill = percent), color=NA) +
  coord_sf(datum=NA) +
  labs(title = "Rate of unsolved murders in Boston by neighborhood",
       subtitle = "Between 2008 and 2018",
       caption = "Source: Washington Post",
       fill = "Percent unsolved") +
  scale_fill_viridis(option="magma", direction=-1) +
  facet_wrap(~victim_race)
```

That's a slick-looking graphic.

These are the results as a data frame if you didn't map it:

```{r dataframe}
points_in %>% 
  group_by(Name, victim_race, disposition) %>%
  summarize(murders=n()) %>% 
  mutate(percent=murders/sum(murders)*100) %>% 
  filter(disposition=="Open/No arrest") 
```


### Add it to an interactive map

One last map.

Let's figure out where unsolved murders occurr in context of Black-majority population neighborhoods.

I've already loaded the Census tracts with the percent of Black population as the data frame **boston_black_perc**. 

Here's how the polygons look:

```{r tracts}
ggplot(boston_black_perc) +
  geom_sf()
``` 

And here's how the data frame looks:

```{r blackpop}
glimpse(boston_black_perc)
```

First, let's create a new data frame that breaks out census tracts where the Black population is the majority.

```{r breakout, exercise=TRUE, warning=F}
black_boston <- boston_black_perc %>% 
  filter(__________ > 50)

black_boston
```

```{r breakout-check}
black_boston <- boston_black_perc %>% 
  filter(black_percent > 50)

black_boston
```


### Final map

Alright, I've set up some colors based on "Closed by arrest" and "Open/No arrest".

```
disp <- colorFactor(c("#ff0000", "#13ED3F"), domain=c("Closed by arrest", "Open/No arrest"))
```

We're going to create a layered interactive map.

First layer is all the Boston census tracts with a minimal border (**boston_black_perc**).

Second layer is the Black only census tracts with a thicker border (**black_boston**).

The third and final layer are circles representing solved and unsolved murders (the colors are based on the code above).

Then, add a legend.

```{r leaflet2, exercise=TRUE, warning=F}
leaflet() %>% addProviderTiles("CartoDB.Positron") %>% 
  setView(-71.087010, 42.316240, zoom = 11) %>% 
    add_______(data=________,
              fillColor = "transparent", 
              fillOpacity = 0.9, 
              weight = 1, 
              smoothFactor = 0.2,
              color="black") %>% 
    add_______(data=________, 
              fillColor = "transparent", 
              fillOpacity = 1, 
              weight = 3, 
              smoothFactor = 0.2,
              color="black") %>% 
    add_______(data=homicides, ~homicides$lon, ~homicides$lat, popup=homicides$_________,
               weight = 3, radius=40, color=~______(disposition), stroke = TRUE,
               fillOpacity = 0.8) %>% 
    add_______("bottomright", colors= c("#ff0000", "#13ED3F"), 
              labels=c("Arrest made/Closed", "Open"), title="Disposition") 
```


```{r leaflet2-check}
leaflet() %>% addProviderTiles("CartoDB.Positron") %>% 
  setView(-71.087010, 42.316240, zoom = 11) %>% 
    addPolygons(data=boston_black_perc,
              fillColor = "transparent", 
              fillOpacity = 0.9, 
              weight = 1, 
              smoothFactor = 0.2,
              color="black") %>% 
    addPolygons(data=black_boston, 
              fillColor = "transparent", 
              fillOpacity = 1, 
              weight = 3, 
              smoothFactor = 0.2,
              color="black") %>% 
    addCircles(data=homicides, ~homicides$lon, ~homicides$lat, popup=homicides$disposition, weight = 3, radius=40, color=~disp(disposition), stroke = TRUE, fillOpacity = 0.8) %>% 
    addLegend("bottomright", colors= c("#ff0000", "#13ED3F"), labels=c("Arrest made/Closed", "Open"), title="Disposition") 
```

Congratulations!

You're done with this chapter!


